---
title: "NYC COVID-19 Test Positivity by Modified Zip Code Tabulation Area (modZCTA)"
author: "Sarah Forrest"
date: '`r format(Sys.time(), "%m-%d-%Y")`'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE, dpi = 300, fig.width = 7)

library(tidyverse)
library(httr)
library(lubridate)
```

# Create Dataset 

Read in and tidy `caserate-by-modzcta.csv`, `testrate-by-modzcta.csv`, `percentpositive-by-modzcta.csv`, and `populationestimates-by-modzcta.csv`:

```{r}
caserate_df = 
  read_csv("data/raw/caserate-by-modzcta.csv") %>%
  janitor::clean_names() %>%
  select(week_ending, caserate_10001:caserate_11697) %>%
  pivot_longer(
    starts_with("caserate"),
    names_to = "modzcta",
    names_prefix = "caserate_", 
    values_to = "caserate_100k"
  )

testrate_df = 
  read_csv("data/raw/testrate-by-modzcta.csv") %>%
  janitor::clean_names() %>%
  select(week_ending, testrate_10001:testrate_11697) %>%
  pivot_longer(
    starts_with("testrate"),
    names_to = "modzcta",
    names_prefix = "testrate_", 
    values_to = "testrate_100k"
  )

pctpos_df =
read_csv("data/raw/percentpositive-by-modzcta.csv") %>%
  janitor::clean_names() %>%
  select(week_ending, pctpos_10001:pctpos_11697) %>%
  pivot_longer(
    starts_with("pctpos"),
    names_to = "modzcta",
    names_prefix = "pctpos_", 
    values_to = "pctpos"
  )

modzcta_pop_df = 
  read_csv("data/raw/populationestimates-by-modzcta.csv") %>% 
  janitor::clean_names() %>%
  select(modzcta, pop_est) %>%
  slice(1:(n() - 1))
```

Merge `caserate-by-modzcta.csv`, `testrate-by-modzcta.csv`, and `percentpositive-by-modzcta.csv` by `modzcta` and `week_ending`:

```{r}
merged_df = 
  list(caserate_df, testrate_df, pctpos_df) %>% 
  reduce(full_join, by = c("week_ending", "modzcta")) %>% 
  mutate(
    modzcta = as.numeric(modzcta),
    week_ending = as.Date(week_ending, format = "%m/%d/%Y"))
```

Merge modZCTA population estimates to the dataframe by `modzcta`:

```{r}
merged_df = full_join(merged_df, modzcta_pop_df)
```

Restrict study period to the the weeks ending August 08, 2020 to October 10, 2020:

```{r}
merged_df = 
  merged_df %>%
  filter(week_ending <= '2020-10-10')
```

Calculate `casecount`, `testcount`, and `testpos` (test positivity) and reorder variables:

```{r}
merged_df = 
  merged_df %>% 
  mutate(
    casecount = caserate_100k * pop_est / 100000,
    testcount = testrate_100k * pop_est / 100000,
    testpos = casecount / testcount * 100) %>%
  select(week_ending, modzcta, pop_est, casecount, caserate_100k, testcount, testrate_100k, pctpos, testpos)
```

Export dataframe as a CSV file: 

```{r}
final_df =
  merged_df %>%
  select(-pctpos)

# write.csv(final_df,"data/casecount-testcount-testpos-by-modzcta.csv", row.names = TRUE)
```

# Create Files for SaTScan Analysis

Case File:

```{r}
case_file = 
  final_df %>% 
  select(modzcta, casecount, week_ending) %>% 
  mutate(casecount = round(casecount, digits = 0))

write.csv(case_file,"data/satscan/case_file.csv", row.names = TRUE)
```

Population File:

```{r}
pop_file = 
  final_df %>% 
  select(modzcta, week_ending, testcount) %>% 
  mutate(testcount = round(testcount, digits = 0))

write.csv(pop_file,"data/satscan/population_file.csv", row.names = TRUE)
```

# Create Files for ArcGIS Emerging Hotspot Analysis

```{r}
ehsa_data = 
  final_df %>% 
  select(modzcta, week_ending, testpos, casecount, testcount) %>% 
  mutate(
    casecount = round(casecount, digits = 0),
    testcount = round(testcount, digits = 0))

write.csv(pop_file,"data/arcgis/ehsa_data.csv", row.names = TRUE)
```

# Exploratory Data Analysis

Average Percent Test Positivity across all ModZCTAs over Study Period:

```{r}
final_df %>% 
  group_by(week_ending) %>%
  summarize(mean_testpos = mean(testpos)) %>%
  ggplot(aes(x = week_ending, y = mean_testpos)) + 
  theme(axis.text.x = element_text(angle = 90,hjust = 1)) + geom_line(show.legend = FALSE) +
  ggtitle("Avg NYC COVID-19 Test % Positivity by Week From 8/2/2020 to 10/10/2020") + 
  labs(
    x = "week ending date",
    y = "mean test % positivity")
```

Percent Test Positivity for by ModZCTA over Study Period:

```{r}
final_df %>% 
  mutate(modzcta = as.character(modzcta)) %>% 
  ggplot(aes(x = week_ending, y = testpos, group = modzcta, color = modzcta)) + 
  theme(axis.text.x = element_text(angle = 90,hjust = 1)) + geom_line(show.legend = FALSE) +
  ggtitle("COVID-19 Test % Positivity by ModZCTA by Week From 8/2/2020 to 10/10/2020") + 
  labs(
    x = "week ending date",
    y = "mean test % positivity")
```

## Percent Test Positivity for Each ModZCTA over Study Period

Create dataframe for use in plots: 

```{r}
plot_df = 
  merged_df %>% 
  rename(
    pctpos_reported = pctpos,
    pctpos_calculated = testpos) %>% 
  pivot_longer(
    starts_with("pctpos"),
    names_to = "group",
    names_prefix = "pctpos_", 
    values_to = "value")
```

Create function to plot the % positive value (from `percentpositive-by-modzcta.csv`) vs. my calculated casecount/testcount*100 value for each modZCTA over the study period:

```{r}
plot_modzcta = function(modzcta_entry) {

plot_df %>% 
  filter(modzcta == modzcta_entry) %>% 
  ggplot(aes(x = week_ending, y = value, group = group, color = group)) + 
  geom_line(alpha = 0.7) +
  ylim(0, 9) +
  labs(
    title = modzcta_entry,
    x = "week ending date",
    y = "test % positivity") +
  scale_color_hue(labels = c("calculated" = "casecount/testcount*100", "reported" = "% positive"))
  
}

modzcta_list = modzcta_pop_df$modzcta

for (i in 1:177) {
  
  plot = plot_modzcta(modzcta_list[[i]])
  print(plot)
  
}
```